---
title: "Developing Data Product - Week 3 Assignment"
author: "Dylan"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(dplyr)
```


## Instruction

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity! 


Since this is only a simple demonstration, we will just use the built-in `mtcars`. It's not fancy but it'll get the job done. Will save the effort for final assignment.

```{r data}
data(mtcars)
mtcars <- mtcars %>%
  mutate(
    cyl = as.factor(cyl),
    vs = as.factor(if_else(vs == 0, "v-shaped", "straight")),
    am = as.factor(if_else(am == 0, "automatic", "manual")),
    gear = as.factor(gear)
  )
glimpse(mtcars)
```

## Simple Scatterplot

First scatterplot: weight v. mile per gallon, color by number of cylinders.

```{r scatt1}
scatt1 <- plot_ly(data = mtcars,
                  x = ~wt, y = ~mpg, color = ~cyl,
                  type = "scatter", mode = "markers") %>%
  layout(title = "Scatterplot 1: weight v. Miles per Gallon", legend = list(title = list(text = "Cylinders")))

scatt1
```


Second scatterplot: horsepower v. qsec, size by displacement.

```{r scatt2}
scatt2 <- plot_ly(data = mtcars,
                  x = ~hp, y = ~qsec, size = ~disp,
                  type = "scatter", mode = "markers") %>%
  layout(title = "Scatterplot 2: Horsepower v. Qsec", legend = list(title = list(text = "Displacement")))

scatt2
```



## Simple Bar Chart

Bar Chart 1: Average MPG by number of Cylinders

```{r bar1}
bar1 <- mtcars %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg),
            avg_disp = mean(disp),
            avg_hp = mean(hp)) %>%
  plot_ly(x = ~cyl, y = ~avg_mpg,
          type = "bar") %>%
  layout(
    title = "Average MPG by Number of Cylinders",
    xaxis = list(title = "Number of Cylinders"),
    yaxis = list(title = "Average MPG")
  )

bar1
```


Bar Chart 2: Average Displacement Volume by number of Cylinders

```{r bar2}
bar2 <- mtcars %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg),
            avg_disp = mean(disp),
            avg_hp = mean(hp)) %>%
  plot_ly(x = ~cyl, y = ~avg_disp,
          type = "bar") %>%
  layout(
    title = "Average DISP by Number of Cylinders",
    xaxis = list(title = "Number of Cylinders"),
    yaxis = list(title = "Average DISP")
  )

bar2
```

