---
title: "Developing Data Product - Week 3 Assignment"
author: "Dylan"
date: "2024-05-03"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plotly)
library(dplyr)
```

```{r data}
data(mtcars)
mtcars <- mtcars %>%
  mutate(
    cyl = as.factor(cyl),
    vs = as.factor(if_else(vs == 0, "v-shaped", "straight")),
    am = as.factor(if_else(am == 0, "automatic", "manual")),
    gear = as.factor(gear)
  )
```

## Background

Hi there! We'll just use the built-in `mtcars` data set that already comes with R.
It's nothing fancy, but it gets the job done! Let's save the effort for the final assignment.

- Hi I am bullet point 1
- I contain no info
- Just thought it's fancy
- Also empty space makes me uncomfortable

## Simple Scatterplot 1

```{r scatt1}
scatt1 <- plot_ly(data = mtcars,
                  x = ~wt, y = ~mpg, color = ~cyl,
                  type = "scatter", mode = "markers") %>%
  layout(title = "Scatterplot 1: weight v. Miles per Gallon", legend = list(title = list(text = "Cylinders")))

scatt1
```

## Simple Scatterplot 2

```{r scatt2, warning=FALSE}
scatt2 <- plot_ly(data = mtcars,
                  x = ~hp, y = ~qsec, size = ~disp,
                  type = "scatter", mode = "markers") %>%
  layout(title = "Scatterplot 2: Horsepower v. Qsec", legend = list(title = list(text = "Displacement")))

scatt2
```

## Simple Bar Chart 1

```{r bar1}
bar1 <- mtcars %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg),
            avg_disp = mean(disp),
            avg_hp = mean(hp)) %>%
  plot_ly(x = ~cyl, y = ~avg_mpg,
          type = "bar") %>%
  layout(
    title = "Average MPG by Number of Cylinders",
    xaxis = list(title = "Number of Cylinders"),
    yaxis = list(title = "Average MPG")
  )

bar1
```

## Simple Bar Chart 2

```{r bar2}
bar2 <- mtcars %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg),
            avg_disp = mean(disp),
            avg_hp = mean(hp)) %>%
  plot_ly(x = ~cyl, y = ~avg_disp,
          type = "bar") %>%
  layout(
    title = "Average DISP by Number of Cylinders",
    xaxis = list(title = "Number of Cylinders"),
    yaxis = list(title = "Average DISP")
  )

bar2
```

## That's it folks!

We're done. Now close the browser and go get some sunlight outside. Or moonlight. Whenever you may be. Idk, don't let me tell you what to do. I am not your boss.
