---
title: "Data Product - Demo Presentation"
author: "Dylan"
date: "2024-04-29"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


## My new slide

What does the stuff say?

**This is double**

*This is single*


### How about H3?

What would it say?

#### This is H4 subheading


## My title

- Bullet 1
- Bullet 2

1. Item a
2. Item b
3. Item c

Let's see what it looks like.


## Formatting Options

*Italicized text*

**Bold text**

`Cold chunks for (i %in% 1:10)`
`second line of code chunk`

```
Blocky code chunk
line2
3
```


## R Code that runs

Take a look at code!
`comment = ""` takes away the comment #

```{r, comment = ""}
head(mtcars)
```


## R code II

Also, by default `echo = false`
So if you want to show the code you'll have to specify.

```{r}
print("Hello Hal.")
```

Or you can remove the first few lines to default to `echo = TRUE`
As well as specifying other options.

## Slide: Plot

```{r message=FALSE}
require(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    xlab = "Weight",
    xlab = "MPG"
  )
```
